---
title: "PS8"
author: "Stros"
date: "12/22/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r , include=TRUE}
library(dplyr)
library(dslabs)
library(purrr)
library(ggplot2)
library(tidyverse)


p <- setwd("C:/Users/Sebastian/Desktop/Humbdoldt/ML but political")
library(AER)
library(ivpack)
library(haven)
library(readr)
library(stargazer)
library(cobalt)
d <-   readRDS ("combined_data.RDS")

head(d)
library(MatchIt)
 


```
#### 1.
```{r , include=TRUE}
summary(lm(data= d, pctVV~ treat))
```
#### The treatment seems to be large with however quite large SE and very very low R squared. On the upside the p-value is very good.

#### 2.
```{r , include=TRUE}

 dm <- d %>% filter(male == 1)
 summary(lm(data = dm, pctVV~ treat))
 
  dw <- d %>% filter(male == 0)
 summary(lm(data = dw, pctVV~ treat))
 
 summary(lm(data = d, treat~ male))

```
#### There does seem to be a bias of being selected into the treatment group conditional on being male and the outcomes depending on the sexes singificantly differ. 
Honestly I was not sure what the question asked whether it was the outcomes given the different sexes or the proclivity of being selected into treatment as a male.

#### 3.

```{r , include=TRUE}
 summary(lm(data = d, pctVV~ treat + male))
```
#### The treatment effect is thus a tiny bit smaller, but the SE stays the same.


d_treated <-  d %>% filter(treat == 1)
d_control <-  d %>% filter(treat == 0)

n <-length(d_treated$male)

d_treated <-  d_treated[,"matched_control_unit_pctVV"] <- NA

for (i in n){
  gender_treated <- d_treated$male[i]
  
  d_control <-  d %>% filter(treat == 0)
  ## this line feels needed so that I don't have only the gender in d_control defined by previous iteration
  
   d_control <- d_control %>% filter(gender_treated == d_control$male )
   ## the matching
  
   d_control <- sample_n( d_control, 1)
   
   d_treated[i,"matched_control_unit_pctVV"] <-  d_control$pctVV
   ## maybe here is a failure in the assignment??
 
  
     
}
d_treated$matched_control_unit_pctVV
## checking for the results


#### 4.

```{r , include=TRUE}


d_treated <- d %>% filter(treat==1)
d_controled <- d %>% filter(treat==0)
d_treated["matched_control_unit_pctVV"] <- NA

for (i in 1:dim(d_treated)[1]){
  
  gender_treated <- d_treated[i,] %>% select(male) %>% as.numeric()
  
  matched_unit <- d_controled %>% filter(male==gender_treated) %>% sample_n(1)
  d_treated[i,"matched_control_unit_pctVV"] <- matched_unit %>% select(pctVV) %>% as.numeric()
}






 
```
#### 5. With this matched dataset, we now have balance on gender by definition. To estimate the treatment effect using a difference-in-means calculate the difference in outcomes between your treated and matched control units in the treated units dataset (i.e. between the pctVV and matched_control_unit_pctVV variables). Then average these differences to get an average treatment effect and interpret the result. Also conduct a t-test to get a p-value.


```{r , include=TRUE}

new_d <- d_treated %>% mutate(difference_in_outcomes=pctVV-matched_control_unit_pctVV) %>% summarise(diff_outcome=mean(difference_in_outcomes,na.rm=TRUE))

new_d

 t.test(x = d_treated$pctVV,y = d_treated$matched_control_unit_pctVV)$p.value 
```

### 6.
```{r , include=TRUE}
 d6 <- d_treated %>% select(pctVV, matched_control_unit_pctVV)
d6 <- d6 %>% gather(key = "Treatment",value="pctVV_all")
d6 <- d6 %>% mutate(Treatment = ifelse(Treatment == "pctVV", 1, 0) )
summary(lm(data = d6,  pctVV_all ~ Treatment))



```
It is very similar, albeit we get p-value above 0.05


### 7.
```{r , include=TRUE}
output <- matchit(treat ~ log.valid.votes, data = d)
summary(output)
```
the question's suggested "output$nn" returned "NULL"

### 8.
```{r , include=TRUE}

diff_original <- mean(d_treated$log.valid.votes) - mean(d_controled$log.valid.votes) 
diff_original
diff_matched_c <- match.data(output) %>% filter(treat == 0) 
diff_matched_t <- match.data(output) %>% filter(treat == 1) 

mean(diff_matched_t$log.valid.votes) - 
mean(diff_matched_c$log.valid.votes)
```
 
 
### 9.
```{r , include=TRUE}
d %>% ggplot() + geom_point(aes(x=log.valid.votes,y=pctVV,color=factor(treat)), shape = 2, alpha=0.2) +
  geom_point(data=match.data(output),aes(x=log.valid.votes,y=pctVV,color=factor(treat)))

```
### 10.
```{r , include=TRUE}
 summary(lm(data = match.data(output), pctVV ~ treat))
```
There seems to be a strong impact of 22.6%, but the p-value is too large and the adjusted R-squared is way too small.
### 11. Conduct the same regression as in Q8 but add a control for log.valid.votes, the variable we already matched on. Does this change the estimate of the treatment effect compared to your answer in Q8? Why/why not?
```{r , include=TRUE}
 summary(lm(data = match.data(output), pctVV ~ treat + log.valid.votes))

```
Question 8. was not a regression so I guess it was referring to question 10. Anyway there is basically no difference in the coefficient but there is much better p-value.
### 12.
```{r , include=TRUE}
covars <- c("occBlue.collar", "occEducation", "occGovernment", "occMedia", "occNone", "occOther", "occPolitician", "occWhite.collar", "lat", "long", "ran.prior", "incumbent", "log.valid.votes", "party.prior.pctVV", "prior.pctVV", "elec.year", "match.partyPCB", "match.partyPC.do.B", "match.partyPDT", "match.partyPFL", "match.partyPL", "match.partyPMDB", "match.partyPMN", "match.partyPP", "match.partyPPS", "match.partyPSB", "match.partyPSC", "match.partyPSDB", "match.partyPSDC", "match.partyPSL", "match.partyPT", "match.partyPTB", "match.partyPV", "uf.rs", "uf.sp", "yob", "eduMore.than.Primary..Less.than.Superior", "eduSome.Superior.or.More", "log.total.assets", "pt_pres_1998", "psdb_2000", "hdi_2000", "income_2000", "log.num.apps")

covars_formula <- paste0(covars,collapse=" + ")
matched_data <- matchit(as.formula(paste0("treat~",covars_formula)), data=d,method="nearest")
```
### 13.
```{r , include=TRUE}
treat_sds <- d %>% filter(treat==1) %>% summarise_at(covars,sd)

fulldata_mean_diff <- d %>% group_by(treat) %>% summarise_at(covars,mean) %>% summarise_all(diff) %>% select(-treat)
fulldata_mean_diff_std <- (fulldata_mean_diff/treat_sds)*100 
fulldata_mean_diff_std <- fulldata_mean_diff_std %>% gather()

fulldata_p <- d %>%  summarise_at(covars,funs(t.test(.[treat==0], .[treat==1])$p.value)) %>% gather()

final <- match.data(matched_data)

treat_sds_matched <- final %>% filter(treat==1) %>% summarise_at(covars,sd)

matcheddata_mean_diff <- final %>% group_by(treat) %>% summarise_at(covars,mean) %>% summarise_all(diff) %>% dplyr::select(-treat)
matcheddata_mean_diff_std <- (matcheddata_mean_diff/treat_sds_matched)*100
matcheddata_mean_diff_std <- matcheddata_mean_diff_std %>% gather()

matched_p <- final %>% summarise_at(covars,funs(t.test(.[treat==0], .[treat==1])$p.value)) %>% gather()

fulldata_mean_diff_std %>% left_join(matcheddata_mean_diff_std, by=c("key")) %>% 
  left_join(fulldata_p,by=c("key")) %>%
  left_join(matched_p,by=c("key")) %>%
  dplyr::rename(Before_Mean=value.x, After_Mean=value.y, Before_p=value.x.x, After_p=value.y.y) %>% 
  mutate_at(vars(Before_Mean,After_Mean,Before_p,After_p),funs(round(.,3)))
```
### 14.

```{r , include=TRUE}
lm(pctVV ~ treat, data = final)
```
### 15.
```{r , include=TRUE}


d_15 <- matchit(as.formula(paste0("treat~",covars_formula)), data=d,method="nearest", caliper = 0.01)

df <- match.data(d_15)

lm(pctVV ~ treat, data = df)

```
### 16.
```{r , include=TRUE, cache=TRUE}
library(rgenoud)
library(Matching)
data_16 <- matchit(as.formula(paste0("treat~",covars_formula)), data=d,method="genetic") %>% match.data()
lm(pctVV ~ treat, data = data_16 )
```
### 17.
I dont really know. Any variable that is not included in covars. Perhaps flu wave could influence both the treatment assignment and the pctVV variable.
