---
title: "PS5"
author: "Stros"
date: "11/26/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r , include=TRUE}
library(dplyr)
library(dslabs)
library(purrr)
library(ggplot2)
library(dslabs)
library(tidyverse)
data(heights)
data(murders)
rm(list = ls())
getwd()
p <- setwd("C:/Users/Sebastian/Desktop/Humbdoldt/ML but political")
library(AER)
library(ivpack)
library(haven)
library(readxl)

  library(stargazer)
  filename <- "d2000.csv"

fullpath <- file.path(filename)
fullpath
dir <- system.file(p) 

file.copy(fullpath, filename)
d0 <- read.csv(fullpath)
 
 filename <- "d2004.csv"

fullpath <- file.path(filename)
fullpath
dir <- system.file(p) 

file.copy(fullpath, filename)
d4 <- read.csv(fullpath)


```
### 2.
#### a.

```{r , include=TRUE}
d0 <- d0 %>% filter( NUM_TURNO == 1) %>%
  select(COD_MUN_IBGE, SIGLA_PARTIDO, QTDE_VOTOS) %>%
  group_by(COD_MUN_IBGE)%>%
  mutate(vote_pct_2000 = QTDE_VOTOS/ sum(QTDE_VOTOS), 
         Electorate = sum(QTDE_VOTOS)) 

d0 <-  d0 %>% mutate(Rank = rank(-QTDE_VOTOS))
d0




 
```
#### b.


```{r , include=TRUE}


d0 <- d0 %>% arrange(COD_MUN_IBGE, Rank) %>% 
  mutate(Highest = max(vote_pct_2000) )  %>%
  mutate(Second = nth(vote_pct_2000,2))


                                                       

d0
```
#### c.

```{r , include=TRUE}

d0 <- d0 %>% mutate(win_margin=ifelse(Rank==1,vote_pct_2000-Second,vote_pct_2000-Highest))



```

#### d.
```{r , include=TRUE}

d0 <- d0 %>% mutate(Incumbent = 
                      ifelse(QTDE_VOTOS == Highest, 1, 0))
d0
```
### 3.

```{r , include=TRUE}
d4 <- d4 %>% filter( NUM_TURNO == 1) %>%
  select(COD_MUN_IBGE, SIGLA_PARTIDO, QTDE_VOTOS) %>%
  group_by(COD_MUN_IBGE)%>%
  mutate(vote_pct_2004 = QTDE_VOTOS/ sum(QTDE_VOTOS),
         Electorate = sum(QTDE_VOTOS)) 
d4
```

```{r , include=TRUE}
results1 <- d0 %>% filter( SIGLA_PARTIDO == "PMDB")
results2 <- d4 %>% filter( SIGLA_PARTIDO == "PMDB")

results_PMDB <- results1 %>% left_join (results2, by = c("COD_MUN_IBGE", "SIGLA_PARTIDO"))
results_PMDB
```
#### There are less rows than before. 2,648
### 5.

```{r , include=TRUE}
 lm(results_PMDB$vote_pct_2004 ~ results_PMDB$Incumbent)

```
#### 6.


```{r , include=TRUE}



cutoff<- results_PMDB %>% filter(win_margin<0.05 & win_margin>-0.05)
cutoff

cutoff %>% 
  group_by(Incumbent) %>% 
  summarise(Electorate=mean(Electorate.x)) 




  

```
### 7.

```{r , include=TRUE}
library(rdd)

DCdensity(results_PMDB$win_margin, verbose = TRUE)

```
Honestly, I don't really know what this means except for the p-value which is over 0.05 which leads to failur of rejecting the null hypothesis.

```{r , include=TRUE}
results_PMDB

ggplot() + geom_point(data=results_PMDB,aes(x=win_margin,y=vote_pct_2004))  + geom_smooth(data=results_PMDB[results_PMDB$win_margin<0,],aes(x=win_margin,y=vote_pct_2004),method="lm")  + geom_smooth(data=results_PMDB[results_PMDB$win_margin>0,],aes(x=win_margin,y=vote_pct_2004),method="lm") + geom_vline(aes(xintercept=0),col="red",size=1)

ggplot() + geom_point(data=results_PMDB,aes(x=win_margin,y=vote_pct_2004)) + geom_smooth(data=results_PMDB[results_PMDB$win_margin<0,],aes(x=win_margin,y=vote_pct_2004),method="loess")  + geom_smooth(data=results_PMDB[results_PMDB$win_margin>0,],aes(x=win_margin,y=vote_pct_2004),method="loess") +
geom_vline(aes(xintercept=0),col="red",size=1)

```
#### The cutoff Incumbancy (ie. just winning) predicts greater chance of losing in the next elections.
### 9.
```{r , include=TRUE}


```
### 10. I dont know why the the mean method cannot be applied on the vote_pct_2004. It says it must be numeric or logical and that vote_pct_2004 is not either, but to me it seems like a column of numerics. I thought that the problem is caused by the NAs so I included the na.pass() so that the NAs are skipped but it seems to be of futile.
```{r , include=TRUE}
bandwidth <- 0.03
 
mean1<- results_PMDB %>% filter(win_margin<bandwidth) %>% mean(vote_pct_2004,na.rm=TRUE)
mean2 <- results_PMDB %>% filter(win_margin>bandwidth ) %>% mean(vote_pct_2004, na.rm=TRUE)

diff_means <- abs(mean1-mean2)
```
### 11.

```{r , include=TRUE}
lm(vote_pct_2004~win_margin+Incumbent + I(win_margin^2) + I(win_margin^3),data=results_PMDB)

```
### 12.

```{r , include=TRUE}
 lm(vote_pct_2004~win_margin+Incumbent,data=results_PMDB)
```
```{r , include=TRUE}
bandwidth <- 0.05
 filtered <- results_PMDB %>% filter(win_margin<bandwidth & win_margin>-bandwidth)

 lm(vote_pct_2004~win_margin+Incumbent,data=filtered)
```
There is fairly high negative coefficient for win_margin. This suggests that in the smaller bandwidth being Incumbent predicts worse success
### 14.
```{r , include=TRUE}
REstimate <- RDestimate(vote_pct_2004 ~ win_margin, data=results_PMDB, cutpoint=0)

REstimate$est

REstimate$p

```
There seems to be a consistent negative effect of win_margin on vote success in 2004; however the negative effect slightly fluctuates depending on the bandwidth. The p-values are small enough so that we can reject the null hypothesis.


```{r , include=TRUE}
 lm(vote_pct_2004~win_margin+Incumbent,data=results_PMDB)
```
###  16.

I would mention that Incumbency has a negative effect for future elections; however this is supported only by data close to the win/lose cutoff of the first election. Winning 30% in the first elections is too far from the studied data. So, I cannot say anything to him/her for sure.
